---
title: "Mountain-data-scraping"
author: "Isha Doshi"
date: "2022-10-28"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Parsing the list of mountains from Wikipedia**

**Loaded the Wikipedia list of mountains by height**
```{r}
library(rvest)
library(tidyverse)
page = read_html("https://en.wikipedia.org/wiki/List_of_mountains_by_elevation")
page
```

**Found all the tables there in the html - 9 tables**

```{r}
page %>%
  html_nodes("table") 
```


**Found the table headers, and determined which columns are mountain names, heights, and where are the links to the individual mountain pages.**

```{r}
page %>%
  html_nodes("table") %>%
  html_elements("th") 

page %>%
  html_nodes("table") %>%
  html_elements("td")

# by comparing th and td we can see that column "Mountain" has mountain name, "Metres" and "Feet" have heights and links to the mountain pages are present in "Mountain" column.
```

**Created a data frame that contains names and heights of the mountains above 6800m, and the links to the corresponding Wikipedia pages.**

```{r}
library(dplyr)
library(rvest)
library(tidyverse)
library (plyr)
library(xml2)
tbls_ls <- page %>%
        html_nodes("table") %>%
        .[1:3] %>%
        html_table(fill = TRUE) 
mountainsDf=bind_rows(tbls_ls)

links= page%>%
  html_nodes("table") %>%
  html_nodes("tr") %>%
  html_nodes(xpath="//td[1]//a[1]") %>%
  html_attr("href")%>%
  head(175)

mountainsDf$Metres <- as.numeric(gsub(",","",mountainsDf$Metres))
mountainsDf <- mountainsDf[mountainsDf$Metres>6800,]
final=cbind(mountainsDf,links)
finalDf=final%>%select(Mountain,Metres,Feet, links)

head(finalDf,5)
nrow(finalDf)
```

**Printing a small sample of my data frame to see that it looks reasonable.**

```{r}
head(finalDf,5)
```


**Wrote a function that converts the longitude/latitude string to degrees (positive and negative).**

```{r}

longitudeAndLatitude <- function(ddmmssDD)
{
D <- if(grepl("[WS]", ddmmssDD)) -1 else 1
dms <- strsplit(ddmmssDD, "°|'|″|′|N|E")
if(length(dms[[1]])==4){
dd=as.numeric(dms[[1]][1])
mm=as.numeric(dms[[1]][2])
ss=as.numeric(dms[[1]][3])
return((dd + mm/60 + ss/3600) *D)
}
else {
dd=as.numeric(dms[[1]][1])
mm=as.numeric(dms[[1]][2])
return((dd + mm/60) *D)
}
}
ans=longitudeAndLatitude("76°38′E")
ans
```

**Wrote another function that takes link as an argument and loads the mountain’s html page and extracts latitude and longitude.**

```{r}
getlatlong=function(link){
  
  url=paste("https://en.wikipedia.org",link, sep = "")
  page=try(read_html(url),silent=TRUE)
  if (inherits (page, "try-error"))
  return (NULL)
  
  latitude=page%>%
  html_element(xpath="//span[@class='latitude']")%>%
  html_text()
  
  longitude=page%>%
  html_element(xpath="//span[@class='longitude']")%>%
  html_text()
  return(c(latitude,longitude))
}

```

**Looped over the table of mountains I made above, downloaded the mountain data, and extracted the coordinates. Stored these into the same data frame.**
```{r}
 latitudes=c();
 longitudes=c();
for(i in finalDf$links)
{
  ll=getlatlong(i)
   latitudes<-append(latitudes,(longitudeAndLatitude(ll[1])))
   longitudes<-append(longitudes,(longitudeAndLatitude(ll[2])))
}
finalDf$Latitudes<-latitudes
finalDf$Longitudes<-longitudes
finalDf
```

**Printing a sample of the dataframe and check that it looks good.**

```{r}
head(finalDf,10)
nrow(finalDf)
```


**Plotting all the mountains on a world map. Colored them according to their height.**

```{r}
data <- data.frame (finalDf$Mountain, longitude=finalDf$Longitudes,
latitude=finalDf$Latitudes)

library (ggplot2)

world <- map_data("world")
MountainsMap<-ggplot (world) +
geom_polygon(aes (long, lat, group=group),
col="white", fill="gray") + geom_point (data=data, aes (longitude, latitude, colour=finalDf$Metres)) +
coord_quickmap ()
MountainsMap + scale_color_gradient(low = "blue", high = "red")
```

**Most of the tallest mountains are located near India, Nepal, Tibet, Bhutan, Pakistan and China. There are two near Argentina and Chile.**
